Optimal Task Phasing for End-To-End Latency in Harmonic Systems and Semi-Harmonic Automotive Systems

### 2024-09-26 paper.tex
structure of harmonic section

\newtheorem{thm}{Theorem}%[section]
\newtheorem{prop}[thm]{Proposition}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{obs}[thm]{Observation}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem{notn}[thm]{Notation}

\theoremstyle{remark}
\newtheorem{rmk}[thm]{Remark}
\newtheorem{exmpl}[thm]{Example}
% --

@@

\newtheorem{lemma}{Lemma}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
%\newtheorem{prog}[theorem]{Programming}
\newtheorem{prog}{Programming}

@@

	Naturally, if $\Tbb$ is harmonic, then $\Tbb$ and all its subsets $\Tbb' \subseteq \Tbb$ are also max-harmonic. 
	In particular, $\Tbb_E$ is max-harmonic.

	To develop an optimal task phasing strategy, we first formulate a general lower bound on the end-to-end latency.

	\begin{lemma}
		The end-to-end latency of $E$ under \emph{any} task phasing is lower bounded by $\sum_{i=1}^{n} T_{\tau_i} + \max_{i=1, \dots, n} T_{\tau_i}$.
	\end{lemma}

		\textbf{Summand of Equation~\eqref{eq:summand1}:}
		Equally, this summand can be written as the sum
		$\eqref{eq:summand1} = \sum_{i=p+1}^{n} (\we(\tau_i(j'_i)) - \we(\tau_{i-1}(j'_{i-1})))$.
		We use the property $\we(\tau_{i-1}(j'_{i-1})) \leq \re(\tau_i(j'_i))$ of immediate forward job chains to bound \eqref{eq:summand1} from below by 
		$\eqref{eq:summand1} \geq \sum_{i=p+1}^{n} (\we(\tau_i(j'_i)) - \re(\tau_{i}(j'_{i})))$.
		Since $\we(\tau_i(j'_i)) - \re(\tau_{i}(j'_{i})) = T_{\tau_i}$, we obtain $\eqref{eq:summand1} \geq \sum_{i=p+1}^{n} T_{\tau_i}$.

		\textbf{Summand of Equation~\eqref{eq:summand2}:}
		By definition, $j_p = m$ and $j'_p = m+1$.
		Therefore, $\eqref{eq:summand2} = (\phi_{\tau_p} + (m+1+1) \cdot T_{\tau_p}) - (\phi_{\tau_p} + m \cdot T_{\tau_p}) = 2\cdot T_{\tau_p}$.

		\textbf{Summand of Equation~\eqref{eq:summand3}:}
		This summand is
		$\eqref{eq:summand3} = \sum_{i=1}^{p-1} ( \re(\tau_{i+1}(j_{i+1})) - \re(\tau_i(j_i)) )$.
		We use the property $\re(\tau_{i+1}(j_{i+1})) \geq \we(\tau_i(j_i))$ of immediate backward job chains to achieve
		$\eqref{eq:summand3} \geq \sum_{i=1}^{p-1} ( \we(\tau_i(j_i)) - \re(\tau_i(j_i)) ) = \sum_{i=1}^{p-1} T_{\tau_i}$.

		We conclude $\ell(\pc^p_m) = \eqref{eq:summand1} + \eqref{eq:summand2} + \eqref{eq:summand3} \geq \sum_{i=1}^{n} T_{\tau_i} + T_{\tau_p} = \sum_{i=1}^{n} T_{\tau_i} + \max_{i=1, \dots, n} T_{\tau_i}$.
		Hence, $\lat(E) =\sup_{m\geq W_p} \ell(\pc^p_m) \geq \sum_{i=1}^{n} T_{\tau_i} + \max_{i=1, \dots, n} T_{\tau_i}$ under any task phasing.
		%
		Since the task phasing $\phi_{\tau_i} = \sum_{\xi=1}^{i-1} T_{\tau_\xi}$ achieves the minimal end-to-end latency, as stated in Theorem~\ref{thm:key}, this phasing is optimal.
	\end{proof}

	In the following, we develop a task phasing that ensures that the end-to-end latency reaches this lower bound for max-harmonic tasks.

@@

	\begin{lemma}\label{lem:key_lemma}
		If $\setof{\tau_1, \dots, \tau_n}$ is semi-harmonic (in the sense as described in Section~\ref{sec:system_model}), then 
		\begin{equation}
			\ell(\pc^p_{W_p}) 
			= \ell(\pc^p_{W_p+1}) 
			= \ell(\pc^p_{W_p+2})
			= \dots 
		\end{equation}
		where $\tau_p$ is the task with the highest period, i.e., $T_{\tau_p} = \max_{i=1, \dots, n} T_{\tau_i}$.
	\end{lemma}

@@

		\textbf{Induction step} ($i \mapsto i-1$):
		By definition of immediate backward job chains, 
		$j_{i-1} = \max \set{\xi \in \Nbb}{ \phi_{\tau_{i-1}} + (\xi+1) \cdot T_{\tau_{i-1}} \leq \phi_{\tau_i} + j_i \cdot T_{\tau_{i}}}$.
		Therefore, $j_{i-1}$ can be formulated as 
		$j_{i-1} = \floor{\frac{\phi_{\tau_i} - \phi_{\tau_{i-1}} + j_i \cdot T_{\tau_i}}{T_{\tau_{i-1}}}} -1$.
		Similarly, $\tilde{j}_{i-1}$ can be written as 
		$\tilde{j}_{i-1} = \floor{\frac{\phi_{\tau_i} - \phi_{\tau_{i-1}} + \tilde{j}_i \cdot T_{\tau_i}}{T_{\tau_{i-1}}}} -1$.
		We know by the induction hypothesis, that $\tilde{j}_i = j_i + \frac{T_{\tau_p}}{T_{\tau_i}}$.
		Therefore, 
		\begin{math}
			\tilde{j}_{i-1} 
			= \floor{\frac{\phi_{\tau_i} - \phi_{\tau_{i-1}} + j_i \cdot T_{\tau_i} + T_{\tau_p}}{T_{\tau_{i-1}}}} -1
			= \floor{\frac{\phi_{\tau_i} - \phi_{\tau_{i-1}} + j_i \cdot T_{\tau_i}}{T_{\tau_{i-1}}} + \frac{T_{\tau_p}}{T_{\tau_i}}} -1
			\overset{\text{(a)}}{=} \floor{\frac{\phi_{\tau_i} - \phi_{\tau_{i-1}} + j_i \cdot T_{\tau_i}}{T_{\tau_{i-1}}}} -1 + \frac{T_{\tau_p}}{T_{\tau_i}}
			= j_{i-1} + \frac{T_{\tau_p}}{T_{\tau_i}}
		\end{math}
		% \begin{align}
		% 	\tilde{j}_{i-1} 
		% 	&= \floor{\frac{\phi_{\tau_i} - \phi_{\tau_{i-1}} + j_i \cdot T_{\tau_i} + T_{\tau_p}}{T_{\tau_{i-1}}}} -1
		% 	= \floor{\frac{\phi_{\tau_i} - \phi_{\tau_{i-1}} + j_i \cdot T_{\tau_i}}{T_{\tau_{i-1}}} + \frac{T_{\tau_p}}{T_{\tau_i}}} -1
		% 	\\& \overset{\text{(a)}}{=} \floor{\frac{\phi_{\tau_i} - \phi_{\tau_{i-1}} + j_i \cdot T_{\tau_i}}{T_{\tau_{i-1}}}} -1 + \frac{T_{\tau_p}}{T_{\tau_i}}
		% 	= j_{i-1} + \frac{T_{\tau_p}}{T_{\tau_i}}
		% \end{align}
		where (a) holds because $\frac{T_{\tau_p}}{T_{\tau_i}} \in \Nbb$ by the semi-harmonic property. \qed~Equation~\eqref{eq:induction1}

@@

		Since we have shown that $j'_{i+1} > W_{i+1} \geq 0$, we know that $j'_{i+1}>0$.
		Hence, the maximum in $j'_{i+1} = \max\setof{\ceiling{\frac{\phi_{\tau_i} - \phi_{\tau_{i+1}} + (j'_i+1) \cdot T_{\tau_i}}{T_{\tau_{i+1}}}} , 0}$ always returns the first value. 
		That means that $j'_{i+1}$ can be formulated as
		\begin{math}
			j'_{i+1} = \ceiling{\frac{\phi_{\tau_i} - \phi_{\tau_{i+1}} + (j'_i+1) \cdot T_{\tau_i}}{T_{\tau_{i+1}}}}.
		\end{math}
		Similarly, we show that $\tilde{j}'_{i+1} > W_{i+1}$ and that 
		\begin{math}
			\tilde{j}'_{i+1} = \ceiling{\frac{\phi_{\tau_i} - \phi_{\tau_{i+1}} + (\tilde{j}'_i+1) \cdot T_{\tau_i}}{T_{\tau_{i+1}}}}.
		\end{math}
		%
		We utilize the induction hypothesis $\tilde{j}'_i = j'_i + \frac{T_{\tau_p}}{T_{\tau_i}}$ to obtain
		\begin{math}
			\tilde{j}'_{i+1} 
			= \ceiling{\frac{\phi_{\tau_i} - \phi_{\tau_{i+1}} + (j'_i +1) \cdot T_{\tau_i} + T_{\tau_p}}{T_{\tau_{i+1}}}}
			= \ceiling{\frac{\phi_{\tau_i} - \phi_{\tau_{i+1}} + (j'_i +1) \cdot T_{\tau_i}}{T_{\tau_{i+1}}} + \frac{T_{\tau_p}}{T_{\tau_{i+1}}}}
			= \ceiling{\frac{\phi_{\tau_i} - \phi_{\tau_{i+1}} + (j'_i +1) \cdot T_{\tau_i}}{T_{\tau_{i+1}}}} + \frac{T_{\tau_p}}{T_{\tau_{i+1}}}
			= j'_{i+1} + \frac{T_{\tau_p}}{T_{\tau_{i+1}}}
		\end{math}
		% \begin{align}
		% 	\tilde{j}'_{i+1} 
		% 	&= \ceiling{\frac{\phi_{\tau_i} - \phi_{\tau_{i+1}} + (j'_i +1) \cdot T_{\tau_i} + T_{\tau_p}}{T_{\tau_{i+1}}}}
		% 	= \ceiling{\frac{\phi_{\tau_i} - \phi_{\tau_{i+1}} + (j'_i +1) \cdot T_{\tau_i}}{T_{\tau_{i+1}}} + \frac{T_{\tau_p}}{T_{\tau_{i+1}}}}
		% 	\\&
		% 	= \ceiling{\frac{\phi_{\tau_i} - \phi_{\tau_{i+1}} + (j'_i +1) \cdot T_{\tau_i}}{T_{\tau_{i+1}}}} + \frac{T_{\tau_p}}{T_{\tau_{i+1}}}
		% 	= j'_{i+1} + \frac{T_{\tau_p}}{T_{\tau_{i+1}}}
		% \end{align}
		where $\frac{T_{\tau_p}}{T_{\tau_{i+1}}} \in \Nbb$ because the periods are semi-harmonic.
		\qed~Equation~\eqref{eq:induction2}

@@

	Due to the previous lemma, $\lat(E) = \ell(\pc^p_{W_p})$.
	Hence, to minimize the end-to-end latency, it is sufficient to choose phases which minimize the \emph{first} partitioned job chain. 
	In the following we \textbf{set the task phases to}
	\begin{equation}
		\phi_{\tau_i} = \sum_{\xi=1}^{i-1} T_{\tau_\xi},
	\end{equation}
	as illustrated in Figure~\ref{fig:phasing}.
	\begin{figure}
		\centering
		\footnotesize
		\begin{tikzpicture}[yscale=0.22, xscale=0.38]

@@

			\end{scope}
			\end{tikzpicture}	
			\caption{Phase enforcement. The first $3$-partitioned job chain for $E=(\tau_1 \to \tau_2 \to \dots \to \tau_5)$ are marked in gray.}
			\label{fig:phasing}
	\end{figure}
	From the figure we observe that the read- and write-event of subsequent jobs coincide for the first $p$-partitioned job chain, which minimizes the length of that job chain.
	In the following theorem we calculate the end-to-end latency of the task set with the specified task phases. 
	Afterwards, we discuss the optimality of that task phasing.

	\begin{theorem}[End-to-end latency with task phasing]\label{thm:key}
		If ${\tau_1, \dots, \tau_n}$ is semi-harmonic and the task offsets are chosen as $\phi_{\tau_i} = \sum_{\xi=1}^{i-1} T_{\tau_\xi}$, then 
		\begin{equation}
			\lat(E) = \sum_{i=1}^{n} T_{\tau_i} + \max_{i=1, \dots, n} T_{\tau_i}
		\end{equation}
	\end{theorem}

	\mario{Instead, this theorem should be about that the end-to-end latency becomes optimal}

@@

	Theorem: Phases need to be set at

	Proposition: How big is the gain that we get wrt synchronous?

>>>>

\newtheorem{theorem}{Theorem}%[section]
% \newtheorem{prop}[thm]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
% \newtheorem{obs}[thm]{Observation}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
% \newtheorem{notn}[thm]{Notation}

% \theoremstyle{remark}
% \newtheorem{rmk}[thm]{Remark}
% \newtheorem{exmpl}[thm]{Example}

%\newtheorem{prog}[theorem]{Programming}
% \newtheorem{prog}{Programming}

% \newtheorem{corollary}{Corollary}
% --

@@

@@

	Naturally, if $\Tbb$ is harmonic, then $\Tbb$ and all its subsets $\Tbb' \subseteq \Tbb$, including $\Tbb_E$, are max-harmonic.

@@

	\begin{lemma}\label{lem:key_lemma}
		Let $\tau_p$ be the task with the highest period of $E$, i.e., $T_{\tau_p} = \max_{i=1, \dots, n} T_{\tau_i}$.
		If $\Tbb_E$ is max-harmonic, then 
		\begin{equation}
			\ell(\pc^p_{W_p}) 
			= \ell(\pc^p_{W_p+1}) 
			= \ell(\pc^p_{W_p+2})
			= \dots 
		\end{equation}
		In particular,  
		\begin{math}
			\lat(E) = \ell(\pc^p_{W_p}) 
		\end{math}
		holds.
	\end{lemma}

@@

		\textbf{Induction step} ($i \mapsto i-1$):
		By definition of immediate backward job chains, 
		$j_{i-1} = \max \set{\xi \in \Nbb}{ \phi_{\tau_{i-1}} + (\xi+1) \cdot T_{\tau_{i-1}} \leq \phi_{\tau_i} + j_i \cdot T_{\tau_{i}}}$.
		Therefore, $j_{i-1}$ can be formulated as 
		$j_{i-1} = \floor{\frac{\phi_{\tau_i} - \phi_{\tau_{i-1}} + j_i \cdot T_{\tau_i}}{T_{\tau_{i-1}}}} -1$.
		Similarly, $\tilde{j}_{i-1}$ can be written as 
		$\tilde{j}_{i-1} = \floor{\frac{\phi_{\tau_i} - \phi_{\tau_{i-1}} + \tilde{j}_i \cdot T_{\tau_i}}{T_{\tau_{i-1}}}} -1$.
		We know by the induction hypothesis, that $\tilde{j}_i = j_i + \frac{T_{\tau_p}}{T_{\tau_i}}$.
		Therefore, 
		\begin{math}
			\tilde{j}_{i-1} 
			= \floor{\frac{\phi_{\tau_i} - \phi_{\tau_{i-1}} + j_i \cdot T_{\tau_i} + T_{\tau_p}}{T_{\tau_{i-1}}}} -1
			= \floor{\frac{\phi_{\tau_i} - \phi_{\tau_{i-1}} + j_i \cdot T_{\tau_i}}{T_{\tau_{i-1}}} + \frac{T_{\tau_p}}{T_{\tau_i}}} -1
			\overset{\text{(a)}}{=} \floor{\frac{\phi_{\tau_i} - \phi_{\tau_{i-1}} + j_i \cdot T_{\tau_i}}{T_{\tau_{i-1}}}} -1 + \frac{T_{\tau_p}}{T_{\tau_i}}
			= j_{i-1} + \frac{T_{\tau_p}}{T_{\tau_i}}
		\end{math}
		% \begin{align}
		% 	\tilde{j}_{i-1} 
		% 	&= \floor{\frac{\phi_{\tau_i} - \phi_{\tau_{i-1}} + j_i \cdot T_{\tau_i} + T_{\tau_p}}{T_{\tau_{i-1}}}} -1
		% 	= \floor{\frac{\phi_{\tau_i} - \phi_{\tau_{i-1}} + j_i \cdot T_{\tau_i}}{T_{\tau_{i-1}}} + \frac{T_{\tau_p}}{T_{\tau_i}}} -1
		% 	\\& \overset{\text{(a)}}{=} \floor{\frac{\phi_{\tau_i} - \phi_{\tau_{i-1}} + j_i \cdot T_{\tau_i}}{T_{\tau_{i-1}}}} -1 + \frac{T_{\tau_p}}{T_{\tau_i}}
		% 	= j_{i-1} + \frac{T_{\tau_p}}{T_{\tau_i}}
		% \end{align}
		where (a) holds because $\frac{T_{\tau_p}}{T_{\tau_i}} \in \Nbb$ by the max-harmonic property. \qed~Equation~\eqref{eq:induction1}

@@

		Since we have shown that $j'_{i+1} > W_{i+1} \geq 0$, we know that $j'_{i+1}>0$.
		Hence, the maximum in $j'_{i+1} = \max\setof{\ceiling{\frac{\phi_{\tau_i} - \phi_{\tau_{i+1}} + (j'_i+1) \cdot T_{\tau_i}}{T_{\tau_{i+1}}}} , 0}$ always returns the first value. 
		That means that $j'_{i+1}$ can be formulated as
		\begin{math}
			j'_{i+1} = \ceiling{\frac{\phi_{\tau_i} - \phi_{\tau_{i+1}} + (j'_i+1) \cdot T_{\tau_i}}{T_{\tau_{i+1}}}}.
		\end{math}
		Similarly, we show that $\tilde{j}'_{i+1} > W_{i+1}$ and that 
		\begin{math}
			\tilde{j}'_{i+1} = \ceiling{\frac{\phi_{\tau_i} - \phi_{\tau_{i+1}} + (\tilde{j}'_i+1) \cdot T_{\tau_i}}{T_{\tau_{i+1}}}}.
		\end{math}
		%
		We utilize the induction hypothesis $\tilde{j}'_i = j'_i + \frac{T_{\tau_p}}{T_{\tau_i}}$ to obtain
		\begin{math}
			\tilde{j}'_{i+1} 
			= \ceiling{\frac{\phi_{\tau_i} - \phi_{\tau_{i+1}} + (j'_i +1) \cdot T_{\tau_i} + T_{\tau_p}}{T_{\tau_{i+1}}}}
			= \ceiling{\frac{\phi_{\tau_i} - \phi_{\tau_{i+1}} + (j'_i +1) \cdot T_{\tau_i}}{T_{\tau_{i+1}}} + \frac{T_{\tau_p}}{T_{\tau_{i+1}}}}
			= \ceiling{\frac{\phi_{\tau_i} - \phi_{\tau_{i+1}} + (j'_i +1) \cdot T_{\tau_i}}{T_{\tau_{i+1}}}} + \frac{T_{\tau_p}}{T_{\tau_{i+1}}}
			= j'_{i+1} + \frac{T_{\tau_p}}{T_{\tau_{i+1}}}
		\end{math}
		% \begin{align}
		% 	\tilde{j}'_{i+1} 
		% 	&= \ceiling{\frac{\phi_{\tau_i} - \phi_{\tau_{i+1}} + (j'_i +1) \cdot T_{\tau_i} + T_{\tau_p}}{T_{\tau_{i+1}}}}
		% 	= \ceiling{\frac{\phi_{\tau_i} - \phi_{\tau_{i+1}} + (j'_i +1) \cdot T_{\tau_i}}{T_{\tau_{i+1}}} + \frac{T_{\tau_p}}{T_{\tau_{i+1}}}}
		% 	\\&
		% 	= \ceiling{\frac{\phi_{\tau_i} - \phi_{\tau_{i+1}} + (j'_i +1) \cdot T_{\tau_i}}{T_{\tau_{i+1}}}} + \frac{T_{\tau_p}}{T_{\tau_{i+1}}}
		% 	= j'_{i+1} + \frac{T_{\tau_p}}{T_{\tau_{i+1}}}
		% \end{align}
		where $\frac{T_{\tau_p}}{T_{\tau_{i+1}}} \in \Nbb$ because the periods are max-harmonic.
		\qed~Equation~\eqref{eq:induction2}

@@

	The previous lemma indicates that, to minimize the end-to-end latency, it is sufficient to choose phases which minimize the \emph{first} partitioned job chain.
	We achieve that, by ensuring that there is no slack between read- and write-events of $\pc^p_{W_p}$.
	In particular, we consider the following task phasing.

	\begin{definition}[Suggested Phasing for Max-Harmonic]\label{def:phasing_max_harm}
		For the cause-effect chain $E= (\tau_1 \to\dots\to \tau_n)$, we define a task phasing
		\begin{equation}\label{eq:phasing}
			\phi^{\mathit{MH}}_{\tau_i} = \sum_{\xi=1}^{i-1} T_{\tau_\xi},
		\end{equation}
		for all $\tau_i \in \Tbb_E$.
	\end{definition}

	The phasing is illustrated in Figure~\ref{fig:phasing}.
	\begin{figure}
		\centering
		\footnotesize
		\begin{tikzpicture}[yscale=0.22, xscale=0.38]

@@

			\end{scope}
			\end{tikzpicture}	
			\caption{Phase enforcement. The first $3$-partitioned job chain for $E=(\tau_1 \to \tau_2 \to \dots \to \tau_5)$ are marked in gray.}
			\label{fig:phasing}
	\end{figure}
	Indeed, we observe that the read- and write-event of subsequent jobs coincide for the first $p$-partitioned job chain, which minimizes the length of that job chain.
	Specifically, the end-to-end latency under the suggested task phasing for max-harmonic systems is as follows.

	\begin{theorem}[End-to-end latency with suggested phasing]\label{thm:e2e_with_phasing_harm}
		If $\Tbb_E$ is max-harmonic and the task offsets are chosen as $\phi_{\tau_i} = \phi^{\mathit{MH}}_{\tau_i}$, then 
		\begin{equation}
			\lat(E) = \sum_{i=1}^{n} T_{\tau_i} + \max_{i=1, \dots, n} T_{\tau_i}.
		\end{equation}
	\end{theorem}

@@

	In the following, we show that the suggested phasing is optimal.
	To that end, we formulate a general lower bound on the end-to-end latency.

	\begin{lemma}\label{lem:lower_bound_general}
		The end-to-end latency of $E$ under \emph{any} task phasing is lower bounded by 
		\begin{equation}
			\lat(E) \geq \sum_{i=1}^{n} T_{\tau_i} + \max_{i=1, \dots, n} T_{\tau_i}.
		\end{equation}
	\end{lemma}

		\smallskip
		\textbf{Summand of Equation~\eqref{eq:summand1}:}
		Equally, this summand can be written as the sum
		$\eqref{eq:summand1} = \sum_{i=p+1}^{n} (\we(\tau_i(j'_i)) - \we(\tau_{i-1}(j'_{i-1})))$.
		We use the property $\we(\tau_{i-1}(j'_{i-1})) \leq \re(\tau_i(j'_i))$ of immediate forward job chains to bound \eqref{eq:summand1} from below by 
		$\eqref{eq:summand1} \geq \sum_{i=p+1}^{n} (\we(\tau_i(j'_i)) - \re(\tau_{i}(j'_{i})))$.
		Since $\we(\tau_i(j'_i)) - \re(\tau_{i}(j'_{i})) = T_{\tau_i}$, we obtain $\eqref{eq:summand1} \geq \sum_{i=p+1}^{n} T_{\tau_i}$.

		\smallskip
		\textbf{Summand of Equation~\eqref{eq:summand2}:}
		By definition, $j_p = m$ and $j'_p = m+1$.
		Therefore, $\eqref{eq:summand2} = (\phi_{\tau_p} + (m+1+1) \cdot T_{\tau_p}) - (\phi_{\tau_p} + m \cdot T_{\tau_p}) = 2\cdot T_{\tau_p}$.

		\smallskip
		\textbf{Summand of Equation~\eqref{eq:summand3}:}
		This summand is
		$\eqref{eq:summand3} = \sum_{i=1}^{p-1} ( \re(\tau_{i+1}(j_{i+1})) - \re(\tau_i(j_i)) )$.
		We use the property $\re(\tau_{i+1}(j_{i+1})) \geq \we(\tau_i(j_i))$ of immediate backward job chains to achieve
		$\eqref{eq:summand3} \geq \sum_{i=1}^{p-1} ( \we(\tau_i(j_i)) - \re(\tau_i(j_i)) ) = \sum_{i=1}^{p-1} T_{\tau_i}$.

		\smallskip
		We conclude $\ell(\pc^p_m) = \eqref{eq:summand1} + \eqref{eq:summand2} + \eqref{eq:summand3} \geq \sum_{i=1}^{n} T_{\tau_i} + T_{\tau_p} = \sum_{i=1}^{n} T_{\tau_i} + \max_{i=1, \dots, n} T_{\tau_i}$.
		Hence, $\lat(E) =\sup_{m\geq W_p} \ell(\pc^p_m) \geq \sum_{i=1}^{n} T_{\tau_i} + \max_{i=1, \dots, n} T_{\tau_i}$ under any task phasing.
	\end{proof}

	Since the lower bound coincides with the end-to-end latency for max-harmonic systems with our suggested phasing, our phasing is optimal.

	\begin{theorem}
		The task phasing presented in Equation~\eqref{eq:phasing} is optimal in minimizing the end-to-end latency for max-harmonic systems.
	\end{theorem}

	\begin{proof}
		Since the end-to-end latency under our proposed phasing (Theorem~\ref{thm:e2e_with_phasing_harm}) coincides with the lower bound under any phasing (Lemma~\ref{lem:lower_bound_general}), our task phasing is optimal.
	\end{proof}

	\mario{If we need to fill space we can evaluate how large the gap can become wrt to random phasing and synchronous phasing.}
